---
title: "p8105_hw2_yj2581"
author: "YucongJiang"
date: "2019-10-2"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
```

## Problem 1

#### Read and clean data in three sheets

```{r trash_wheel, warning = FALSE, message = FALSE}

# Read and clean data in Mr. Trash Wheel
mr_trash_wheel <- read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
                             sheet = "Mr. Trash Wheel", 
                             range = "A2:N408") %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(sports_balls = as.integer(round(sports_balls)))

# Read and clean data in 2017 and 2018 Precipitation
precipitation_2017 <- read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
                                 sheet = "2017 Precipitation",
                                 skip = 1) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2017)

precipitation_2018 <- read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
                                 sheet = "2018 Precipitation",
                                 skip = 1) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2018)

# Join the data in the latter two sheets
precipitation_combine <- full_join(precipitation_2017, precipitation_2018) %>%
  mutate(month = as.character(month.name[month])) %>%
  select(year, month, total)

# Show the result
precipitation_combine
```

#### Describe the data

* The number of observations in `mr_trash_wheel` is `r nrow(mr_trash_wheel)`.
* The number of observations in `precipitation_combine` is `r nrow(precipitation_combine)`.
* The key variables in `mr_trash_wheel` are `r colnames(mr_trash_wheel)`.
* The key variables in `precipitation_combine` are `r colnames(precipitation_combine)`.
* The total precipitation in 2018 is `r sum(pull(precipitation_2018, total))`.
* The median number of sports balls in a dumpster in 2017 is `r median(pull(filter(mr_trash_wheel, year == 2017), sports_balls))`.

## Problem 2

#### Read and clean data in the three files

```{r fivethirtyeight, warning = FALSE, message = FALSE}

# Read and clean data in pols-month.csv
pols_month <- read_csv("data/pols-month.csv") %>%
  separate(mon, c("year", "month", "day"), "-", convert = TRUE) %>%
  mutate(month = as.character(month.name[month])) %>%
  mutate(president = if_else(prez_gop == 1, "gop", "dem")) %>%
  select(-day, -prez_gop, -prez_dem)

# Read and clean data in snp.csv
snp <- read_csv("data/snp.csv") %>%
  separate(date, c("month", "day", "year"), "/", convert = TRUE) %>%
  arrange(year, month) %>%
  mutate(month = as.character(month.name[month])) %>%
  select(year, month, close)

#Read and clean data in unemployment.csv
unemployment <- read_csv("data/unemployment.csv") %>%
  rename(
    year = Year,
    January = Jan,
    February = Feb,
    March = Mar,
    April = Apr,
    June = Jun,
    July = Jul,
    August = Aug,
    September = Sep,
    October = Oct,
    November = Nov,
    December = Dec) %>%
  pivot_longer(
    January:December,
    names_to = "month",
    values_to = "unemployment_percentage") %>%
  drop_na(unemployment_percentage)

# Merge the snp into pols, then merge the unemployment into the result
merge_table <- full_join(pols_month, snp, by = c("year", "month")) %>%
  full_join(unemployment, by = c("year", "month"))

```

#### Describe the data

* The dataset `pols_month` contains `r nrow(pols_month)` observations from `r pull(pols_month, month)[1]`, `r pull(pols_month, year)[1]` to `r pull(pols_month, month)[nrow(pols_month)]`, `r pull(pols_month, year)[nrow(pols_month)]`.
* The dataset `snp` contains `r nrow(snp)` observations from `r pull(snp, month)[1]`, `r pull(snp, year)[1]` to `r pull(snp, month)[nrow(snp)]`, `r pull(snp, year)[nrow(snp)]`.
* The dataset `unemployment` contains `r nrow(unemployment)` observations from `r pull(unemployment, month)[1]`, `r pull(unemployment, year)[1]` to `r pull(unemployment, month)[nrow(unemployment)]`, `r pull(unemployment, year)[nrow(unemployment)]`.
* The variables in `pols_month` include the number of senators, governors and representatives who are democratic or republican at any given time, together with whether the president is republican or democratic.
* The variables in `snp` include the closing values of the S&P stock index on one day of the associated month.
* The variables in `unemployment` include the percentage of unemployment in the associated month.
* **Notice**: The resulting dataset `merge_table` is the **full join** of the three dataset above.
* The dataset `merge_table` has `r nrow(merge_table)` rows and `r ncol(merge_table)` columns.
* The dataset `merge_table` contains observations ranging from `r pull(merge_table, month)[1]`, `r pull(merge_table, year)[1]` to `r pull(merge_table, month)[nrow(merge_table)]`, `r pull(merge_table, year)[nrow(merge_table)]`.
* The key variables in `merge_table` are `r colnames(merge_table)`.

## Problem 3

#### Load and tidy the data

* There are abbreviate expressions in the column 'Ethnicity', replace them with full expressions.
* The characters in the column 'Child's First Name' become all upper case in the latter rows, unify the case.
* Remove duplicated rows.

```{r popular_baby_names}

# Read and clean the data based on the above instructions
popular_baby_names <- read.csv("data/Popular_Baby_Names.csv") %>%
  janitor::clean_names() %>%
  mutate(
    ethnicity = replace(ethnicity, ethnicity == "ASIAN AND PACI", "ASIAN AND PACIFIC ISLANDER"),
    ethnicity = replace(ethnicity, ethnicity == "BLACK NON HISP", "BLACK NON HISPANIC"),
    ethnicity = replace(ethnicity, ethnicity == "WHITE NON HISP", "WHITE NON HISPANIC"),
    ethnicity = as.factor(as.character(ethnicity)),
    child_s_first_name = str_to_title(child_s_first_name)
  ) %>%
  distinct()
```

#### Produce tables

```{r table_data}

# Split and tidy necessary data from the original dataset based on the instructions
olivia_popularity <- filter(popular_baby_names, child_s_first_name == "Olivia" & gender == "FEMALE") %>%
  select(ethnicity, year_of_birth, rank) %>%
  rename(Ethnicity = ethnicity) %>%
  arrange(year_of_birth) %>%
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "rank"
  )

most_popular_boyname <- filter(popular_baby_names, rank == 1 & gender == "MALE") %>%
  select(ethnicity, year_of_birth, child_s_first_name) %>%
  rename(Ethnicity = ethnicity) %>%
  arrange(year_of_birth) %>%
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "child_s_first_name"
  )
```


#### Rank of 'Olivia' as a Female Baby Name Over Time
```{r show_table_1}

knitr::kable(olivia_popularity, caption = "Rank of 'Olivia' as a Female Baby Name Over Time")
```

#### Most Popular Boy Name Over Time
```{r show_table_2}

knitr::kable(most_popular_boyname, caption = "Most Popular Boy Name Over Time")
```

#### Produce a scatter plot

```{r show_plot}
# Split necessary data from the original dataset based on the instructions
count_vs_rank <- filter(popular_baby_names, gender == "MALE" & ethnicity == "WHITE NON HISPANIC" & year_of_birth == 2016) %>%
  select(count, rank)

# Show the scatter plot, add labels.
ggplot(count_vs_rank, aes(x = rank, y = count)) + 
  geom_point() +
  labs(
    title = "Count of Name in Male White Non-hispanic Children Born in 2016",
    x = "The Rank in Popularity",
    y = "The Count of Name",
    caption = "Data from NYC Open data"
  )
```


